---
title: "DM-15"
author: "Gabriele and Dorian"
format: html
---

# Settings
## Here i am

```{r}
here :: i_am("DM-15.Rproj")
```
# Packages

```{r}
library(here)
library(vroom)
library(tidyverse)
```

##Data Import
```{r}
EDUSPEND<-vroom(here("PROJET","EDUCATION DATA SETS","Education spending (OECD).csv"))
EDUTAIN<-vroom(here("PROJET","EDUCATION DATA SETS","Educational attainment, at least completed short-cycle tertiary, population 25+, total (World Bank).csv"))
GDPMFP<-vroom(here("PROJET","GDP & MFP DATASETS","GDP, MFP and GDP per hour worked (OECD).csv"))
GROSPEND<-vroom(here("PROJET","R&D DATA SETS","Gross domestic spending in R&D, Researchers per countries & Tridic Patent families (OECD).csv"))
```

## Github link

https://github.com/gabrielepiantoni/DM-15.git

## Sources' Links

*OECD: https://data-explorer.oecd.org*

In order to use to get the data from OECD Explore you have to select from the panel on the left the following;
- Frequency & Time: From 1914 to 2022
- Country: Select All the Countries
- Indicator:  Education spending, Gross domestic product (GDP), GDP per hour worked, Multifactor productivity, Researchers, Gross domestic spending on R&D, Triadic patent

*World Bank: https://data.worldbank.org/indicator/SE.TER.CUAT.ST.ZS?end=2023&start=2023&view=bar*

Click on CSV in the download section of the bottom right corner of the page