---
title: "DM-15"
author: "Gabriele and Dorian"
format: html
---

# Settings
## Here i am

```{r}
#| message: false
here :: i_am("DM-15.Rproj")
```
# Packages

```{r}
#| message: false
library(here)
library(vroom)
library(tidyverse)
library(knitr)
```

##Data Import
```{r}
#| message: false
EDUSPEND<-vroom(here("PROJET","EDUCATION DATA SETS","Education spending (OECD).csv"))
EDUTAIN<-vroom(here("PROJET","EDUCATION DATA SETS","Educational attainment, at least completed short-cycle tertiary, population 25+, total (World Bank).csv"))
GDPMFP<-vroom(here("PROJET","GDP & MFP DATASETS","GDP, MFP and GDP per hour worked (OECD).csv"))
GROSPEND<-vroom(here("PROJET","R&D DATA SETS","Gross domestic spending in R&D, Researchers per countries & Tridic Patent families (OECD).csv"))
```

## Github link

https://github.com/gabrielepiantoni/DM-15.git

# Sources' Links

## OECD

- [OECD Data Explorer](https://data-explorer.oecd.org)

  To access the data from **OECD Explore**, you will need to select the following options from the panel on the left:
  - **Frequency & Time**: From 1914 to 2022
  - **Country**: Select all available countries
  - **Indicator**: 
    - Education Spending
    - Gross Domestic Product (GDP)
    - GDP per Hour Worked
    - Multifactor Productivity
    - Researchers
    - Gross Domestic Spending on R&D
    - Triadic Patent

## World Bank

- [World Bank Indicator for Educational Attainment](https://databank.worldbank.org/home.aspx)

  To access the data from **World Bank Databank**, you will need to select the following options from the panel on the left:
  - **Database**: World Developments Indicators
  - **Country**: Select all available countries
  - **Series**
    - Educational attainment, at least completed short-cycle tertiary, population 25+, total (%)   
  - **Time**: From 2005 to 2022
 
To download the data, click on Download options then **CSV**.

# Datasets and Sources

## Datasets

### Education Data Sets

- **Education Spending** (source: OECD Data Explorer2; 2005-2020):
  Expenditure in education - divided into primary, secondary and tertiary - in both %GDP and US$/Student ratio.

- **Educational attainment**, at least completed short-cycle tertiary, population 25+, total (source: World Bank; 2005-2022):
  Percentage of adults over 25 who have completed a first cycle of tertiary education.

### GDP and MFP Datasets

- **Gross Domestic Product** (source: OECD Data Explorer; 1970-2022):
  Gross Domestic Product in US$ and US$/capita ratio.

- **MFP** (source: OECD Data Explorer; 1985-2022):
  Multifactor productivity (i.e. TFP) expressed in annual growth rates (%) and current year value compared to 2015.

- **GDP per hour worked** (source: OECD Data Explorer; 1970-2022):
  Ratio of GDP to hours worked in both US$ and current year value relative to 2015.

### R&D Dataset

- **Gross Domestic Spending on R&D** (source: OECD Data Explorer; 1981-2022):
  Gross domestic spending on R&D measured in Million US dollar and percentage of GDP.

- **Researchers per countries** (source: OECD Data Explorer; 1981-2022):
  Number of researchers per country, count per head and ratio per 1000 people employed.

- **Triadic Patent families** (source: OECD Data Explorer; 1981-2022):
  Triadic patent families are a set of patents filed at three of these major patent offices: the European Patent Office (EPO), the Japan Patent Office (JPO), and the United States Patent and Trademark Office (USPTO). Triadic patent family counts are attributed to the country of residence of the inventor and to the date when the patent was first registered.
  
## Sources 

### The Organisation for Economic Co-operation and Development (OECD)

The **Organisation for Economic Co-operation and Development (OECD)** is an international organization composed of 38 member countries, established in 1961 to promote economic growth, trade, and global development. Its primary goal is to foster policies that improve the economic and social well-being of people worldwide.

#### Activities in Data Collection and Display

##### Data Collection

- The OECD collects a wide range of data from member and non-member countries on topics such as **economic performance**, **education**, **labor markets**, **trade**, **health**, **energy**, and **environmental sustainability**.
- The organization relies on **standardized methodologies** to ensure data comparability across countries.

##### Analysis and Reporting

- The OECD analyzes the collected data to identify trends, develop **policy recommendations**, and support evidence-based decision-making.
- It publishes its findings in **reports**, **working papers**, and annual flagship publications such as:
  - **Economic Outlook**
  - **Education at a Glance**
  - **Better Life Index**

##### Interactive Platforms

- The OECD provides tools and platforms like **OECD.Stat** and the **OECD Data Portal**, which allow users to access, visualize, and download data in **interactive formats**.
- These platforms enable policymakers, researchers, and the public to explore **cross-country comparisons** and detailed statistical insights.

##### Policy Development

- The organization uses data to guide **policy frameworks** and support collaboration among countries on pressing global issues such as **climate change**, **inequality**, and **digital transformation**.

#### Summary

In summary, the **OECD** plays a critical role in collecting high-quality, comparable data and leveraging it to drive global economic and social improvements through research, policy recommendations, and interactive dissemination tools.


### The World Bank

The **World Bank** is an international financial institution that provides **loans, grants, and technical expertise** to developing countries with the goal of reducing poverty and promoting sustainable economic development. Established in 1944, it is part of the **World Bank Group** and comprises two main entities: the **International Bank for Reconstruction and Development (IBRD)** and the **International Development Association (IDA)**.

#### Activities in Data Collection and Display

##### Data Collection

- The World Bank gathers **extensive global development data** on topics like **poverty**, **education**, **health**, and **economic performance**.
- Its data informs progress tracking for global initiatives such as the **United Nations' Sustainable Development Goals (SDGs)**.

##### Analysis and Reporting

- The World Bank produces analytical resources such as:
  - **World Development Report**
  - **Doing Business Index**
  - **Poverty and Shared Prosperity Series**
- These reports provide **critical insights** into development trends and challenges.

##### Open Data Initiatives

- The World Bank's **Open Data platform** makes its development data freely available to the public.
- The platform offers tools like **dashboards** and **visualizations** for easier exploration and use.

##### Policy and Project Support

- Data is used to design **evidence-based projects and policies** tailored to the unique needs of developing countries.
- This ensures **impactful and efficient interventions**.

#### Summary

In summary, the **World Bank** leverages its extensive data resources to support global development goals, enabling **informed decision-making** and driving efforts to reduce **poverty** and improve **living standards**.

# Data formating and normalization
  This part precedes any further work as, to allow us to show the stylized facts from which we derive our question of interest, we need formatted data that enable us to perform calculations and visualization.
  We proceed as follow :
  
  We clean the data raw data from irrelevant information in contains (as "STRUCTURE_ID", an OECD data classification that doesn't add any information to the data)
  
  For the ease of manipulation, we split the data into different tables, each one displaying one indicator and its different measures  

  In order to work with the data and be able to use tydir, we need to reformat the tables obtained from the OECD and World Bank into tables uniformly formatted as follow : 
  - One variable per column
  - One row per observation
  
  

## Selecting variables

```{r}
EDUSPEND <- EDUSPEND|>select(Pays, Indicateur, Sujet, Mesure, TIME_PERIOD, OBS_VALUE)
EDUTAIN <- EDUTAIN|>select(-c(`Country Code`, `Series Name`, `Series Code`))
GDPMFP <- GDPMFP |> select(Pays, Indicateur, Sujet, Mesure, TIME_PERIOD, OBS_VALUE)
GROSPEND <- GROSPEND |> select(Pays, Indicateur, Sujet, Mesure, TIME_PERIOD, OBS_VALUE)
``` 

## Spliting tables and pivoting


### GDPMFP

```{r}
GDP<-GDPMFP |> filter(Indicateur=="Produit intérieur brut (PIB)") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
MFP <- GDPMFP|>filter(Indicateur=="Productivité multifactorielle")|>pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
PRODUCTIVITY <- GDPMFP |> 
  filter(Indicateur=="PIB par heure travaillée")|>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
```

### EDUSPEND

```{r}
TOTALEDUSPEND <- EDUSPEND|>filter(Indicateur=="Dépenses d'éducation")|>pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
```

### GROSPEND

```{r}

print("Original column names:")
print(colnames(EDUTAIN))

colnames(EDUTAIN) <- gsub("\\s\\[YR\\d{4}\\]", "", colnames(EDUTAIN))
#ChatGPT generated

```


```{r}
TRIAD <- GROSPEND |> 
  filter(Indicateur=="Familles de brevets triadiques") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

STATERSC <- GROSPEND |> 
  filter(Indicateur=="Chercheurs du secteur de l’État") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

RSC <- GROSPEND |> 
  filter(Indicateur=="Chercheurs") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

SPENDRD <- GROSPEND |> 
  filter(Indicateur=="Dépenses intérieures brutes de R-D") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

EDUTAIN <- EDUTAIN |>
  pivot_longer(cols = -`Country Name`,
               names_to = "Year",
               values_to = "Value" 
               )

```


# Data description 
## Synthesis

```{r}
datasets <- list(
  GDP = GDP,
  MFP = MFP,
  PRODUCTIVITY = PRODUCTIVITY,
  RSC = RSC,
  SPENDRD = SPENDRD,
  STATERSC = STATERSC,
  TOTALEDUSPEND = TOTALEDUSPEND,
  TRIAD = TRIAD,
  EDUTAIN = EDUTAIN
)

summary_table <- data.frame(
  Rows = sapply(datasets, nrow),
  Columns = sapply(datasets, ncol)
)

kable(summary_table, format = "markdown", col.names = c("Dataset", "Rows", "Columns"))
```

# Synthetic project Outline

*Objective*

The objective of this project is to study the correlation between the components that influence the TFP1 (Total Factor Productivity) and GDP (Gross Domestic Product) growth between different countries over the last twenty years.

More in detail, economic theory has long emphasised the centrality of TFP growth for the long-term expansion of the economy. Taking this as our starting point, we found it interesting to investigate whether and how some of the elements considered to drive TFP growth have shaped the performance of the economy over the past two decades.

*Project Structure*

To pursue our objective, we decided to structure our project as follows.

First, for the sake of completeness, by referring to the data contained in the ‘GDP DataSet’ we will assess the long-standing theorised correlation between TFP and GDP.

Subsequently, using the Education and Research datasets, we will highlight whether and what relationships exist between the evidence in the datasets and TFP’s trends.



