---
title: "DM-15-FinalPhase"
author: 
- "Dorian Nadal", 
- "Gabriele Piantoni"
format: pdf
---

```{r}
#Setting: Here i am
#| message: false
here :: i_am("DM-15.Rproj")
```

```{r}
#Setting: Packages
#| message: false
library(here)
library(vroom)
library(tidyverse)
library(knitr)
```

```{r}
#Setting: Data Import
#| message: false
EDUSPEND<-vroom(here("PROJET","EDUCATION DATA SETS","Education spending (OECD).csv"))
EDUTAIN<-vroom(here("PROJET","EDUCATION DATA SETS","Educational attainment, at least completed short-cycle tertiary, population 25+, total (World Bank).csv"))
GDPMFP<-vroom(here("PROJET","GDP & MFP DATASETS","GDP, MFP and GDP per hour worked (OECD).csv"))
GROSPEND<-vroom(here("PROJET","R&D DATA SETS","Gross domestic spending in R&D, Researchers per countries & Tridic Patent families (OECD).csv"))
```


# Introduction

  This paper investigate the relationship between research, education and TFP growth. 
  We start by showing that TFP is worth of interest as it is one of the major drivers of economic growth. That demonstrated, we concentrate on our main goal, determining whether or not research and education trigger TFP growth.

# Research question presentation 

  We here point out that countries have experienced growth over the last decades then we use the augmented version of the neoclassical growth model of Solow and Swan (1956) to show that TFP is one of the major drivers of growth, hence justifying the interest of our question.

## Stylized facts



  



# Data sets descriptions

# Data analysis

# Conclusion
