---
title: "DM-15-FinalPhase"
author: 
- "Dorian Nadal"
- "Gabriele Piantoni"
format: pdf
execute:
  echo: false      
  warning: false  
  message: false   
---
---
title: "brutta"
format: pdf
---

---
title: "DM-15-FinalPhase"
author: 
- "Dorian Nadal"
- "Gabriele Piantoni"
format: pdf
execute:
  echo: false      
  warning: false  
  message: false   
---

```{r}
#Setting: Here i am
#| message: false
here :: i_am("DM-15.Rproj")
```

```{r}
#Setting: Packages
#| message: false
library(here)
library(vroom)
library(tidyverse)
library(knitr)
library(broom)
library(knitr)
library(patchwork)
```

```{r}
#Setting: Data Import
#| message: false
EDUSPEND<-vroom(here("PROJET","EDUCATION DATA SETS","Education spending (OECD).csv"))
EDUTAIN<-vroom(here("PROJET","EDUCATION DATA SETS","Educational attainment, at least completed short-cycle tertiary, population 25+, total (World Bank).csv"))
GDPMFP<-vroom(here("PROJET","GDP & MFP DATASETS","GDP, MFP and GDP per hour worked (OECD).csv"))
GROSPEND<-vroom(here("PROJET","R&D DATA SETS","Gross domestic spending in R&D, Researchers per countries & Tridic Patent families (OECD).csv"))
CPI <-vroom(here("PROJET","CPI.csv"))
L<-vroom(here("PROJET","L.csv"))
K<-vroom(here("PROJET","K.csv"))
```
```{r}
#Setting: Here i am
#| message: false
here :: i_am("DM-15.Rproj")
```

```{r}
#Setting: Packages
#| message: false
library(here)
library(vroom)
library(tidyverse)
library(knitr)
library(broom)
library(knitr)
library(patchwork)
```

```{r}
#Setting: Data Import
#| message: false
EDUSPEND<-vroom(here("PROJET","EDUCATION DATA SETS","Education spending (OECD).csv"))
EDUTAIN<-vroom(here("PROJET","EDUCATION DATA SETS","Educational attainment, at least completed short-cycle tertiary, population 25+, total (World Bank).csv"))
GDPMFP<-vroom(here("PROJET","GDP & MFP DATASETS","GDP, MFP and GDP per hour worked (OECD).csv"))
GROSPEND<-vroom(here("PROJET","R&D DATA SETS","Gross domestic spending in R&D, Researchers per countries & Tridic Patent families (OECD).csv"))
CPI <-vroom(here("PROJET","CPI.csv"))
L<-vroom(here("PROJET","L.csv"))
K<-vroom(here("PROJET","K.csv"))
```
```{r}
# Data cleaning, formating : clean the raw data from irrelevant information in contains
EDUSPEND <- EDUSPEND|>select(Pays, Indicateur, Sujet, Mesure, TIME_PERIOD, OBS_VALUE)
EDUTAIN <- EDUTAIN|>select(-c(`Country Code`, `Series Name`, `Series Code`))
GDPMFP <- GDPMFP |> select(Pays, Indicateur, Sujet, Mesure, TIME_PERIOD, OBS_VALUE)
GROSPEND <- GROSPEND |> select(Pays, Indicateur, Sujet, Mesure, TIME_PERIOD, OBS_VALUE)
```

```{r}
# Data cleaning, formating : split the data into different tables, each one displaying one indicator and its different measures
GDP<-GDPMFP |> filter(Indicateur=="Produit intérieur brut (PIB)") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
MFP <- GDPMFP|>filter(Indicateur=="Productivité multifactorielle")|>pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
PRODUCTIVITY <- GDPMFP |> 
  filter(Indicateur=="PIB par heure travaillée")|>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
```

```{r}
# Data cleaning, formating : same as previous chunk
TRIAD <- GROSPEND |> 
  filter(Indicateur=="Familles de brevets triadiques") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

STATERSC <- GROSPEND |> 
  filter(Indicateur=="Chercheurs du secteur de l’État") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

RSC <- GROSPEND |> 
  filter(Indicateur=="Chercheurs") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

SPENDRD <- GROSPEND |> 
  filter(Indicateur=="Dépenses intérieures brutes de R-D") |>
  pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)

EDUTAIN <- EDUTAIN |>
  pivot_longer(cols = -`Country Name`,
               names_to = "Year",
               values_to = "Value" 
               )
```


```{r}
# Data cleaning, formating : setting TOTALEDUSPEND i the right format
TOTALEDUSPEND <- EDUSPEND|>filter(Indicateur=="Dépenses d'éducation")|>pivot_wider(
  names_from = "Mesure",
  values_from = "OBS_VALUE"
)
```

```{r}
# Data cleaning, formating : removing the code of each TIME_PERIOD to allow for datamanipulation
colnames(EDUTAIN) <- gsub("\\s\\[YR\\d{4}\\]", "", colnames(EDUTAIN))
#ChatGPT generated
```
```{r}
#filtering CPI
CPI <-CPI |>
  filter(Sujet=="Total") |>
  select(Pays,TIME_PERIOD,OBS_VALUE)
```

```{r}
# renaming column
CPI<-CPI |>
  rename(CPI = OBS_VALUE)
```

```{r}
#joining GDP and CPI data sets by two columns (Year, Pays)
GDP <-GDP |> 
  full_join(CPI, by=c("TIME_PERIOD","Pays"))
```

```{r}
# Computing realGDP
GDP <-GDP |>
  mutate(RealGDP=((`Millions de dollars US`/CPI)*100))
```

```{r}
# Computing realGDP per capita
GDP <-GDP |>
  mutate(RealGDPperCapita=((`Dollars US/habitant`/CPI)*100))
```

```{r}
# removing years after covid crisis not to have the catch up effect 
GDP <- GDP |>
  filter(TIME_PERIOD<="2020")
```

```{r}
#creating categroies of countries
#excluding Israël because of aberrant values
#excluding Russia because of aberrant values
#excluding Colombia because of aberrant values

GDP <- GDP |>
  mutate(`Euro Area and GB` = ifelse(Pays %in% c("France", "Ex-République fédérale d'Allemagne", "Finlande", "Lettonie", 
  "Slovénie", "Irlande", "Autriche", "Malte", "Italie", "Chypre", "Croatie", 
  "Portugal", "Pays-Bas", "Grèce", "République slovaque", "Luxembourg", 
  "Belgique", "Lituanie", "Estonie", "Espagne","Royaume-Uni" ), "YES", "NO"))
         
GDP <- GDP |>
  mutate(AsianTigers = ifelse(Pays %in% c("Corée", "Singapour"), "YES", "NO"))

GDP <- GDP |>
  mutate(EtatsUnis = ifelse(Pays %in% c("États-Unis"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(Europe = ifelse(Pays %in% c("Royaume-Uni", "France", "Allemagne", "Norvège", "Danemark", "Nouvelle-Zélande", "Suisse", "Finlande", "Suède", "Pays-Bas", "Autriche", "Irlande", "Tchéquie", "Slovénie", "Pologne","Lettonie", "Lituanie", "Estonie", "Bulgarie", "Roumanie", "Croatie", "Portugal", "Italie", "Grèce", "Espagne", "Belgique", "Luxembourg", "Serbie", "Albanie", "Macédoine du Nord"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(CentralEurope = ifelse(Pays %in% c("Allemagne", "Autriche", "Pologne", "Tchéquie", 
                    "Slovénie", "Slovaquie", "Hongrie", "France"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(SouthernEurope = ifelse(Pays %in% c("Portugal", "Espagne", "Italie", "Grèce", 
                     "Croatie", "Malte"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(EasternEurope = ifelse(Pays %in% c("Lettonie", "Lituanie", "Estonie"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(NorthernEurope = ifelse(Pays %in% c("Royaume-Uni", "Norvège", "Danemark", 
                     "Suède", "Finlande", "Pays-Bas", "Irlande"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(China = ifelse(Pays %in% c("Chine (République populaire de)"), "YES", "NO"))
```

```{r}
#creating binary variables to indicate if a country belongs to a category
# List of YES/NO variables
# computing average GDP (variables starting with v)
yes_no_variables <- c( 
                      "AsianTigers", 
                      "Europe",
                      "CentralEurope", 
                      "SouthernEurope", 
                      "EasternEurope", 
                      "NorthernEurope", 
                      "EtatsUnis",
                      "China",
                      "Euro Area and GB")

for (var in yes_no_variables) {
  GDP <- GDP |>
    group_by(TIME_PERIOD) |>
    mutate(!!paste0("v", var) := ifelse(!!sym(var) == "YES", 
                                        mean(RealGDP [!!sym(var) == "YES"], na.rm = TRUE), 
                                        NA)) |>
    ungroup()
}
```

```{r}
# computing average GDP per capita
# variables starting with c
for (var in yes_no_variables) {
  GDP <- GDP |>
    group_by(TIME_PERIOD) |>
    mutate(!!paste0("c", var) := ifelse(!!sym(var) == "YES", 
                                        mean(RealGDPperCapita [!!sym(var) == "YES"], na.rm = TRUE), 
                                        NA)) |>
    ungroup()
}
```


```{r}
#| output: false
#Graphs: Real GDP per capita - Euro Area and GB, ASIANTIGERS, China, united states
graph1 <-GDP |>
  pivot_longer(cols = c("cEtatsUnis", 
                   "cAsianTigers", 
                   "cChina",
                   "cEuro Area and GB"),
               names_to = "Region",
               values_to="avgGDP"
               ) |>
  filter(!is.na(avgGDP)) |>
  filter(TIME_PERIOD >=1995) |>
  ggplot(aes(y=avgGDP, x=TIME_PERIOD, color=Region)) +
  geom_line()
```

```{r}
#| output: false
#Graphs: Real GDP - EUROPE
graph2<-GDP |>
  pivot_longer(cols = c("cEurope", 
                      "cCentralEurope", 
                      "cSouthernEurope", 
                      "cEasternEurope", 
                      "cNorthernEurope"),
               names_to = "Region",
               values_to="avgGDP"
               ) |>
  filter(!is.na(avgGDP)) |>
  filter (TIME_PERIOD>=1995) |>
  ggplot(aes(y=avgGDP, x=TIME_PERIOD, color=Region)) +
  geom_line()
```

# Research question presentation

## Introduction

 In recent decades, most of the countries has achieved sustained positive growth, leading them to a marked improvement in their quality of life.
That said, recent years have shown, especially in the more developed economies, a slow down in this trend. To date, the consensus seems to agree that the deflation of the explosive growth that persisted for most of the 20th century is to be attributed to a less sustained growth in technical change which, according to theory, is strongly linked to the amount of resources devoted to research and education.
Well, given the relevance of the topic, we thought it would be interesting to try to determine if the data are coherent with the existence of a causal relationship between the level of research, education and the growth in technical change exists.

## Stylized facts: the relarionship within groeth rate of GDP and TFP

### The performance of world economies in recent decades: a positive scenario

As mentioned above, the economic development of several countries has undergone positive growth in recent decades. In the following graphs (*Graph 1* and *Graph 2*), this phenomenon is highlighted. 

With the aim of obtaining a more coherent representation of the economic development of a selected group of countries and to express more adequately the level of welfare, the graphs below displays **real GDP per capita** - obtaining applying the inflation (CPI) to the nominalk GDP - over the period between 1995 and 2020.

More in detail,

 - *Graph 1* contains data for the USA, Asian Tigers (North Korea and Singapore), China and EuroArea (2023).
 - *Graph 2* contains data for European countries by geographical location
```{r}
#| fig.width: 12
#| fig.height: 4
graph1 + graph2 +
  plot_layout(widths = c(10, 10))
```

```{r}
#Computing the growth rate of GDP
GDP <- GDP |>
  group_by(Pays) |>
arrange(TIME_PERIOD)|> mutate(growth_rateY = ((RealGDP- lag(RealGDP)) / lag(RealGDP)) * 100) |>
  arrange(Pays,TIME_PERIOD)
```

```{r}
#| output: false
#Graphs: Growth_rate GDP - Germany, France, US and South Corea
graph3 <-GDP |>
  filter(Pays=="France"|Pays== "États-Unis"|Pays== "Allemagne"|Pays== "Corée") |>
  ggplot(aes(x=TIME_PERIOD, y=growth_rateY, color=Pays)) +
  geom_smooth(method="lm", se=FALSE)

```

```{r}
#| output: false
##Graphs: TFP growth - Germany, France, US and South Corea
graph4<-MFP|>
   filter(Pays=="France"|Pays== "Chine (République populaire de)"|Pays== "États-Unis"|Pays== "Allemagne"|Pays== "Corée") |>
  ggplot(aes(x=TIME_PERIOD, y=`Taux de croissance annuel (%)`, color=Pays)) +
  geom_smooth(method="lm", se=FALSE)
```
### Not all that glitters is gold: the slowdown in economic growth

Although growth in recent decades has remained positive, it has been declining in several world economies. *Graph 3* gives an initial and rough idea of this phenomenon.
Looking again at some of the countries included in the sample used in *Graph 1* and *Graph 2*, *Graph 3* shows a fall in the **growth rate of real GDP** in countries of global significance, such as the USA, Germany, France and South Korea.

### *Could technical change (TFP) be the culprit in the growth slowdown?*

Economic literature has always theorised the centrality of technical change within economic expansion.  *Graph 4*, together with the already presented *Graph 3*, attempts to give an initial empirical insight into this relationship.
*Graph* 4, which considers the same countries used in the previous chart, highlights the development of TFP growth between 1970 and 2020. As can easily be seen, the graphical relationship with *Graph 4* is evident, suggesting a possible **causal relationship between TFP and GDP growth**.
```{r}
#| fig.width: 12
#| fig.height: 4
graph3+graph4
```

```{r}
#formatting L
L<-L |>
  select(TIME_PERIOD,Pays,Indicateur, OBS_VALUE )
L<-L |>
  pivot_wider(names_from = Indicateur,
              values_from = OBS_VALUE)
```

```{r}
#Computing total hours worked
L<-L |>
  mutate(L = `Population active`*`Heures travaillées`)
```

```{r}
# Computing the growth rate of L
L<-L |>
  group_by(Pays) |>
arrange(TIME_PERIOD)|> mutate(growth_rateL = ((L- lag(L)) / lag(L)) * 100) |>
  arrange(Pays,TIME_PERIOD)
```

```{r}
# Cleaning K
K<-K |> 
  select(Pays, TIME_PERIOD, Mesure, OBS_VALUE) |>
  filter(Mesure=="Taux de croissance annuel (%)") |>
  pivot_wider(names_from=Mesure, 
              values_from=OBS_VALUE) |>
  arrange(Pays, TIME_PERIOD) |>
  rename(growth_rateK= `Taux de croissance annuel (%)` )
```

```{r}
# Creating the dataset for growth decomposition
Dec <- GDP |>
  select(Pays, TIME_PERIOD, growth_rateY)
```

```{r}
# Creating the dataset for growth decomposition
Dec<-Dec |>
  full_join(K, Dec, by=c("TIME_PERIOD", "Pays"))
```

```{r}
# Creating the dataset for growth decomposition
Dec <- Dec |>
  full_join(L, Dec, by=c("TIME_PERIOD", "Pays")) |>
  select(, Pays, TIME_PERIOD, growth_rateY, growth_rateK, growth_rateL)
```

```{r}
# Creating the dataset for growth decomposition
MFP <- MFP |>
  rename(growth_rateA=`Taux de croissance annuel (%)`)
```

```{r}
# Creating the dataset for growth decomposition
Dec <-Dec |>
  full_join(MFP, Dec, by=c("TIME_PERIOD", "Pays")) |>
  select(Pays, TIME_PERIOD, growth_rateY, growth_rateK, growth_rateL, growth_rateA)
```

```{r}
#| output: false
# Creating the dataset for growth decomposition : checking the accuracy
Dec |>
  mutate(check=growth_rateY-(0.33*growth_rateK+0.66*(growth_rateL+ growth_rateA))) |>
  filter(Pays=="France" | Pays=="Allemagne"| Pays=="Corée")
```
### Technical change and economic growth: a causal relationship

Although the result of the previous section is interesting, its statistical relevance is far from being achieved. For this reason, in the current section we will draw more statistically relevant conclusions.

Considering the typical neoclassical production function
$$
Y = (AL)^{1−\alpha}K^{\alpha}
$$
, output growth (i.e. GDP) - taking logs and derivatives - can be decomposed into the following components: capital growth ($g_K$), labour force growth ($g_L$) and TFP growth ($g_A$). Consequently, the benchmark formula becomes
$$
g_Y = \alpha g_K + (1 - \alpha)(g_A + g_L)
$$
sing $\alpha$=1/3, as it is the commonly used value for this parameter and this value tends to remain the same over time (Kaldor fact: the share of capital in production tends to be constant).
At this point, it is crucial to demonstrate the empirical validity of this formula by referring to the period between 1990 and 2020.
*Figure 5*, which considers the USA, France and Germany, decomposes the growth that occurred over the 30-year period, considering both variables endogenous and exogenous to the model. As can be seen from the representation, the explanatory variables largely justify the economic growth trend.
```{r}
# France
# Solow-Swan model
Solow <- Dec |>
  filter(Pays == "France") |>
  arrange(TIME_PERIOD) |>
  mutate(
    growth_rateY = (100 + growth_rateY) / 100,
    growth_rateK = (100 + growth_rateK) / 100,
    growth_rateL = (100 + growth_rateL) / 100,
    growth_rateA = (100 + growth_rateA) / 100
  ) |>
  filter(TIME_PERIOD >= 1990 & TIME_PERIOD <= 2015) |>
  mutate(
    growth_rate_totalL = 0.66*prod(growth_rateL),
    growth_rate_totalY = prod(growth_rateY),
    growth_rate_totalK = 0.33*prod(growth_rateK),
    growth_rate_totalA = 0.66*prod(growth_rateA)
  ) |>
  mutate(others = growth_rate_totalY - growth_rate_totalK - growth_rate_totalL - growth_rate_totalA)|>
  select(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others, growth_rate_totalY)
```

```{r}
#| output: false
# France 
Solow1 <- Solow |>
  pivot_longer(
    cols = c(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others),
    names_to = "Component",
    values_to = "Value"
  )

# Bar chart decomposition
graph5 <- ggplot(Solow1, aes(x = "", y = Value, fill = Component)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Growth rate decomposition (France, Germany,USA, 1990-2015)",
    x = NULL,
    y = "Growth Rate Contribution",
    fill = "Component"
  ) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
```


```{r}
# Germany
# Solow-Swan model
Solow2 <- Dec |>
  filter(Pays == "Allemagne") |>
  arrange(TIME_PERIOD) |>
  mutate(
    growth_rateY = (100 + growth_rateY) / 100,
    growth_rateK = (100 + growth_rateK) / 100,
    growth_rateL = (100 + growth_rateL) / 100,
    growth_rateA = (100 + growth_rateA) / 100
  ) |>
  filter(TIME_PERIOD >= 1990 & TIME_PERIOD <= 2015) |>
  mutate(
    growth_rate_totalL = 0.66*prod(growth_rateL),
    growth_rate_totalY = prod(growth_rateY),
    growth_rate_totalK = 0.33*prod(growth_rateK),
    growth_rate_totalA = 0.66*prod(growth_rateA)
  ) |>
  mutate(others = growth_rate_totalY - growth_rate_totalK - growth_rate_totalL - growth_rate_totalA)|>
  select(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others, growth_rate_totalY)
```

```{r}
#| output: false
# Germany 
Solow3 <- Solow2 |>
  pivot_longer(
    cols = c(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others),
    names_to = "Component",
    values_to = "Value"
  )

# Bar chart decomposition
graph6<-ggplot(Solow3, aes(x = "", y = Value, fill = Component)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r}
# USA
# Solow-Swan model
Solow4 <- Dec |>
  filter(Pays == "États-Unis") |>
  arrange(TIME_PERIOD) |>
  mutate(
    growth_rateY = (100 + growth_rateY) / 100,
    growth_rateK = (100 + growth_rateK) / 100,
    growth_rateL = (100 + growth_rateL) / 100,
    growth_rateA = (100 + growth_rateA) / 100
  ) |>
  filter(TIME_PERIOD >= 1990 & TIME_PERIOD <= 2015) |>
  mutate(
    growth_rate_totalL = 0.66*prod(growth_rateL),
    growth_rate_totalY = prod(growth_rateY),
    growth_rate_totalK = 0.33*prod(growth_rateK),
    growth_rate_totalA = 0.66*prod(growth_rateA)
  ) |>
  mutate(others = growth_rate_totalY - growth_rate_totalK - growth_rate_totalL - growth_rate_totalA)|>
  select(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others, growth_rate_totalY)
```

```{r}
#| output: false
# USA 
Solow5 <- Solow4 |>
  pivot_longer(
    cols = c(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others),
    names_to = "Component",
    values_to = "Value"
  )

# Bar chart decomposition
graph7 <-ggplot(Solow5, aes(x = "", y = Value, fill = Component)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r}
#| fig.width: 12
#| fig.height: 4
graph5+graph6+graph7
```
Once the reliability of the model has been proven, the causality of the relationship between TFP and GDP growth can be demonstrated using **linear regressions** where $g_Y$ (i.e. dependent variable) is demonstrated through the components (i.e. independent variables) $g_L$, $g_K$ and $g_A$.

In the first instance, the results of an all-inclusive linear regression are demonstrated in *Table 1*. As shown, statistical significance levels (ie. p value < 0.5) are met and an interesting coefficient is attributed to the $g_A$ component, emphasising the intensity of the causal relationship between TFP and GDP.

```{r}
#linear model : regression of L, K, A on Y
summary(lm(formula =growth_rateY~growth_rateK+growth_rateL+growth_rateA, data=na.exclude(Dec))) |>
  tidy()|>
  mutate(p.value=as.character(p.value))|>
  kable( 
      caption = "Summary of Linear Model Coefficients", 
      col.names = c("Term", "Estimate", "Std. Error", "t value", "p value"))
```
On the other hand, *Table 2* shows a  $R^2$ and an adjusted $R^2$ of 0.32. Nevertheless, considering that the aim of our model is not to make forecasts, we can be satisfied with such a value.
```{r}
lm(formula =growth_rateY~growth_rateK+growth_rateL+growth_rateA, data=na.exclude(Dec))|>
summary() |>
  (\(x) data.frame(
    Statistic = c("N", "R2", "Adjusted R2"),
    Value = c(x$df[1] + x$df[2], x$r.squared, x$adj.r.squared)
  ))() |>
  kable(
    caption = "Model Summary Statistics",
    col.names = c("Statistic", "Value")
  )

```


Moreover, *Table 3* shows the results of a linear regression controlling for individual countries. Again, the significance levels for the above components are statistically acceptable and the coefficients tend to the values previously obtained and the causality between GDP and TFP is confirmed. In addition, the coefficient for the binary variable associated with different countries are globally not significant, that is, differences across countries do not add much information about the relation between TFP growth and GDP growth.


```{r}
#linear model : regression of L, K, A on Y

summary(lm(formula =growth_rateY~growth_rateK+growth_rateL+growth_rateA+Pays, data=na.exclude(Dec))) |>
  tidy()|>
  mutate(p.value=as.character(p.value))|>
  kable( 
      caption = "Summary of Linear Model Coefficients", 
      col.names = c("Term", "Estimate", "Std. Error", "t value", "p value"))
```
On the other hand, *Table 4* shows a  $R^2$ of 0.39 and an adjusted $R^2$ of 0.37. Nevertheless, considering that the aim of our model is not to make forecasts, we can be satisfied with such a value.
```{r}
lm(formula = growth_rateY ~ growth_rateK + growth_rateL + growth_rateA + Pays, data = na.exclude(Dec)) |>
summary() |>
  (\(x) data.frame(
    Statistic = c("N", "R2", "Adjusted R2"),
    Value = c(x$df[1] + x$df[2], x$r.squared, x$adj.r.squared)
  ))() |>
  kable(
    caption = "Model Summary Statistics",
    col.names = c("Statistic", "Value")
  )

```
Considering these evidence and the theoretical framework provided, we can believe, with confidence, in the existence of a strong causal relationship between TFP growth and GDP growth.

# Data sets descriptions

# Data analysis

# Conclusion

#OLD PART
  We here point out that countries have experienced growth over the last decades then we use the augmented version of the neoclassical growth model of Solow and Swan (1956) to show that TFP growth causes economic growth, hence justifying the consistence of our question.

We use CPI(Consumption price index) from the OECD to compute the real GDP and GDP per capita. In order to get rid of the effect of inflation in the GDP growth that we compute herafter

```{r}
#filtering CPI
CPI <-CPI |>
  filter(Sujet=="Total") |>
  select(Pays,TIME_PERIOD,OBS_VALUE)
```

```{r}
# renaming column
CPI<-CPI |>
  rename(CPI = OBS_VALUE)
```

```{r}
#joining GDP and CPI data sets by two columns (Year, Pays)
GDP <-GDP |> 
  full_join(CPI, by=c("TIME_PERIOD","Pays"))
```

```{r}
# Computing realGDP
GDP <-GDP |>
  mutate(RealGDP=((`Millions de dollars US`/CPI)*100))
```

```{r}
# Computing realGDP per capita
GDP <-GDP |>
  mutate(RealGDPperCapita=((`Dollars US/habitant`/CPI)*100))
```

```{r}
# removing years after covid crisis not to have the catch up effect 
GDP <- GDP |>
  filter(TIME_PERIOD<="2020")
```

We create categories of countries to simplify the presentation and compute average growth rates (justification to do !!)

```{r}
#creating categroies of countries
#excluding Israël because of aberrant values
#excluding Russia because of aberrant values
#excluding Colombia because of aberrant values

GDP <- GDP |>
  mutate(`Euro Area and GB` = ifelse(Pays %in% c("France", "Ex-République fédérale d'Allemagne", "Finlande", "Lettonie", 
  "Slovénie", "Irlande", "Autriche", "Malte", "Italie", "Chypre", "Croatie", 
  "Portugal", "Pays-Bas", "Grèce", "République slovaque", "Luxembourg", 
  "Belgique", "Lituanie", "Estonie", "Espagne","Royaume-Uni" ), "YES", "NO"))
         
GDP <- GDP |>
  mutate(AsianTigers = ifelse(Pays %in% c("Corée", "Singapour"), "YES", "NO"))

GDP <- GDP |>
  mutate(EtatsUnis = ifelse(Pays %in% c("États-Unis"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(Europe = ifelse(Pays %in% c("Royaume-Uni", "France", "Allemagne", "Norvège", "Danemark", "Nouvelle-Zélande", "Suisse", "Finlande", "Suède", "Pays-Bas", "Autriche", "Irlande", "Tchéquie", "Slovénie", "Pologne","Lettonie", "Lituanie", "Estonie", "Bulgarie", "Roumanie", "Croatie", "Portugal", "Italie", "Grèce", "Espagne", "Belgique", "Luxembourg", "Serbie", "Albanie", "Macédoine du Nord"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(CentralEurope = ifelse(Pays %in% c("Allemagne", "Autriche", "Pologne", "Tchéquie", 
                    "Slovénie", "Slovaquie", "Hongrie", "France"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(SouthernEurope = ifelse(Pays %in% c("Portugal", "Espagne", "Italie", "Grèce", 
                     "Croatie", "Malte"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(EasternEurope = ifelse(Pays %in% c("Lettonie", "Lituanie", "Estonie"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(NorthernEurope = ifelse(Pays %in% c("Royaume-Uni", "Norvège", "Danemark", 
                     "Suède", "Finlande", "Pays-Bas", "Irlande"), "YES", "NO"))
         
GDP <- GDP |>
  mutate(China = ifelse(Pays %in% c("Chine (République populaire de)"), "YES", "NO"))
```

```{r}
#creating binary variables to indicate if a country belongs to a category
# List of YES/NO variables
# computing average GDP (variables starting with v)
yes_no_variables <- c( 
                      "AsianTigers", 
                      "Europe",
                      "CentralEurope", 
                      "SouthernEurope", 
                      "EasternEurope", 
                      "NorthernEurope", 
                      "EtatsUnis",
                      "China",
                      "Euro Area and GB")

for (var in yes_no_variables) {
  GDP <- GDP |>
    group_by(TIME_PERIOD) |>
    mutate(!!paste0("v", var) := ifelse(!!sym(var) == "YES", 
                                        mean(RealGDP [!!sym(var) == "YES"], na.rm = TRUE), 
                                        NA)) |>
    ungroup()
}
```

```{r}
# computing average GDP per capita
# variables starting with c
for (var in yes_no_variables) {
  GDP <- GDP |>
    group_by(TIME_PERIOD) |>
    mutate(!!paste0("c", var) := ifelse(!!sym(var) == "YES", 
                                        mean(RealGDPperCapita [!!sym(var) == "YES"], na.rm = TRUE), 
                                        NA)) |>
    ungroup()
}
```


```{r}
#| output: false
#Graphs: Real GDP per capita - Euro Area and GB, ASIANTIGERS, China, united states
graph1 <-GDP |>
  pivot_longer(cols = c("cEtatsUnis", 
                   "cAsianTigers", 
                   "cChina",
                   "cEuro Area and GB"),
               names_to = "Region",
               values_to="avgGDP"
               ) |>
  filter(!is.na(avgGDP)) |>
  filter(TIME_PERIOD >=1995) |>
  ggplot(aes(y=avgGDP, x=TIME_PERIOD, color=Region)) +
  geom_line()
```

```{r}
#| output: false
#Graphs: Real GDP - EUROPE
graph2<-GDP |>
  pivot_longer(cols = c("cEurope", 
                      "cCentralEurope", 
                      "cSouthernEurope", 
                      "cEasternEurope", 
                      "cNorthernEurope"),
               names_to = "Region",
               values_to="avgGDP"
               ) |>
  filter(!is.na(avgGDP)) |>
  filter (TIME_PERIOD>=1995) |>
  ggplot(aes(y=avgGDP, x=TIME_PERIOD, color=Region)) +
  geom_line()
```
In this part, we show that countries have experienced growth over the past decades, and that this growth is due (that is, is linkled in a causal way)to TFP growth.
We rely on the model of Solow and Swan (1956), who created a growth model that explains economic growth by accumulation of capital , the increase in the *labour force* (maybe is better to say labour in general) and the technological process, which is measdured by the change in TFP. 
To do so, we use data from the OECD. We abstracted the nominal GDP of *number* countries, which we *space* into real GDP *by* deflating by the level of the price, using 2015 as the base year (that is, price level is *one* with the consumption price index (CPI)). We then computed the growth rate of the real GDP for our computations. For capital accumulation. we used the growth rate of the GFCF (in real terms) given by OECD and for the labour force increase we computed, using the hours worked/worker per year and the labour force at disposal in each year, the total number of hours worked every year in each country and, then, its growth rate.

Figure 1 and Figure 2 point out that countries have experienced over the past decades, Both show the evolution of GDP per capita from 1995 to 2020 (data contained *...*). We used GDP per capita for the sake of clarity and for it is a better indicator of welfare; from now, we will only use GDP.
We group countries with similars characteristics to show that growth is a general fact. The Asian Tigers have all experienced all experienced very*space* growth during the second half of the XXi century; the countries that have adopted the euro (2023) *space*

Figure 3 and 4 displaies respectively the evpolution of the GDP growth rate and the one of TFP for Korea (Asian Tigers), France, Germany and USA (no data for China for TFP)The data visualized has suggested a strong correlation, especially for Korea, between TFP and GDP growth.

We know *space* a theoretical framework, affected by empirical evidence, that established a causal relationship between TFP growth and GDP growth.


Figure 5 shows that the data fits very well with the model prediction: the contribution of factors that are not considered is very small compared to the other parts.
We decomposed growth according to the model growth decomposition equation, using $\alpha$=1/3, as it is the commonly used value for this parameter and this value tends to remain the same over time (Kaldor fact: the share of capital in production tend to be constant)
Table 1 and 2 support pur proposition. When we compute a regression to explain g_Y with g_L, g_K and g_A as explaneatpry variable, we find a very high *space* for all parameters and consideratively high coefficent for g_A.
Even when controlling for countries in table 2, we found similar coefficients still with very high significance levels.
We notice that coefficient for the binary variable *space* for countries are globally not significant, that is, differences across coutries do not add much information about the relation between TFP growth and GDP growth.
Considering these evidence and the theoretical framework provided, we can believe, with confidence, in the existence of a strong causal relationship between TFP growth and GDP growth.


# Solow-Swan Model (1956)

The Solow model is based on two simple equations: one production function, one capital accumulation equation. It relies on strong assumptions:

1. Constant return to scale  
2. Decreasing marginal returns of inputs  
3. Constant and exogenous saving rate in the economy  

## Production Function

$$
Y = F(K, L)
$$

$$
Y = (AL)^{1−\alpha}K^{\alpha}
$$

Where:  
-  Y : production (GDP)  
-  K : stock of capital  
-  L : labor force  
-  A : level of technology  

## Capital Accumulation Equation

$$
\frac{dK}{dt} = sF(K, L) − \delta K
$$

Where:  
- s: saving rate (also, net investment rate)  
- $\delta$ : capital depreciation rate  

## Model Predictions

- Each economy converges toward a long-term steady-state path, where new investment just makes up for the loss in capital due to depreciation.  
- The steady states can differ from one country to another.  
- The further away the economy is from its steady state, the higher its growth rate.  

## Long-Term Implications

On the long term, only improvements in technology and the growth of the population lead to growth in production.

Taking logs and derivatives, we find:

$$
g_Y = \alpha g_K + (1 - \alpha)(g_A + g_L)
$$



```{r}
#| fig.width: 12
#| fig.height: 4
graph1 + graph2 +
  plot_layout(widths = c(10, 10))
```

#TFP and GDP growth

Putting face to face TFP and GDP growt-->exist a correlation between the two
```{r}
#Computing the growth rate of GDP
GDP <- GDP |>
  group_by(Pays) |>
arrange(TIME_PERIOD)|> mutate(growth_rateY = ((RealGDP- lag(RealGDP)) / lag(RealGDP)) * 100) |>
  arrange(Pays,TIME_PERIOD)
```

```{r}
#| output: false
#Graphs: Growth_rate GDP - Germany, France, US and South Corea
graph3 <-GDP |>
  filter(Pays=="France"|Pays== "États-Unis"|Pays== "Allemagne"|Pays== "Corée") |>
  ggplot(aes(x=TIME_PERIOD, y=growth_rateY, color=Pays)) +
  geom_smooth(method="lm", se=FALSE)

```

```{r}
#| output: false
##Graphs: TFP growth - Germany, France, US and South Corea
graph4<-MFP|>
   filter(Pays=="France"|Pays== "Chine (République populaire de)"|Pays== "États-Unis"|Pays== "Allemagne"|Pays== "Corée") |>
  ggplot(aes(x=TIME_PERIOD, y=`Taux de croissance annuel (%)`, color=Pays)) +
  geom_smooth(method="lm", se=FALSE)
```

```{r}
#| fig.width: 12
#| fig.height: 4
graph3+graph4
```

# Growth decomposition and regression of TFP and GDP

```{r}
#formatting L
L<-L |>
  select(TIME_PERIOD,Pays,Indicateur, OBS_VALUE )
L<-L |>
  pivot_wider(names_from = Indicateur,
              values_from = OBS_VALUE)
```

```{r}
#Computing total hours worked
L<-L |>
  mutate(L = `Population active`*`Heures travaillées`)
```

```{r}
# Computing the growth rate of L
L<-L |>
  group_by(Pays) |>
arrange(TIME_PERIOD)|> mutate(growth_rateL = ((L- lag(L)) / lag(L)) * 100) |>
  arrange(Pays,TIME_PERIOD)
```

```{r}
# Cleaning K
K<-K |> 
  select(Pays, TIME_PERIOD, Mesure, OBS_VALUE) |>
  filter(Mesure=="Taux de croissance annuel (%)") |>
  pivot_wider(names_from=Mesure, 
              values_from=OBS_VALUE) |>
  arrange(Pays, TIME_PERIOD) |>
  rename(growth_rateK= `Taux de croissance annuel (%)` )
```

Creating a data set for the growth decomposition
```{r}
# Creating the dataset for growth decomposition
Dec <- GDP |>
  select(Pays, TIME_PERIOD, growth_rateY)
```

```{r}
# Creating the dataset for growth decomposition
Dec<-Dec |>
  full_join(K, Dec, by=c("TIME_PERIOD", "Pays"))
```

```{r}
# Creating the dataset for growth decomposition
Dec <- Dec |>
  full_join(L, Dec, by=c("TIME_PERIOD", "Pays")) |>
  select(, Pays, TIME_PERIOD, growth_rateY, growth_rateK, growth_rateL)
```

```{r}
# Creating the dataset for growth decomposition
MFP <- MFP |>
  rename(growth_rateA=`Taux de croissance annuel (%)`)
```

```{r}
# Creating the dataset for growth decomposition
Dec <-Dec |>
  full_join(MFP, Dec, by=c("TIME_PERIOD", "Pays")) |>
  select(Pays, TIME_PERIOD, growth_rateY, growth_rateK, growth_rateL, growth_rateA)
```

```{r}
#| output: false
# Creating the dataset for growth decomposition : checking the accuracy
Dec |>
  mutate(check=growth_rateY-(0.33*growth_rateK+0.66*(growth_rateL+ growth_rateA))) |>
  filter(Pays=="France" | Pays=="Allemagne"| Pays=="Corée")
```

```{r}
#linear model : regression of L, K, A on Y
summary(lm(formula =growth_rateY~growth_rateK+growth_rateL+growth_rateA, data=na.exclude(Dec))) |>
  tidy()|>
  mutate(p.value=as.character(p.value))|>
  kable( 
      caption = "Summary of Linear Model Coefficients", 
      col.names = c("Term", "Estimate", "Std. Error", "t value", "p value"))
```
```{r}
lm(formula =growth_rateY~growth_rateK+growth_rateL+growth_rateA, data=na.exclude(Dec))|>
summary() |>
  (\(x) data.frame(
    Statistic = c("N", "R2", "Adjusted R2"),
    Value = c(x$df[1] + x$df[2], x$r.squared, x$adj.r.squared)
  ))() |>
  kable(
    caption = "Model Summary Statistics",
    col.names = c("Statistic", "Value")
  )

```


```{r}
#linear model : regression of L, K, A on Y

summary(lm(formula =growth_rateY~growth_rateK+growth_rateL+growth_rateA+Pays, data=na.exclude(Dec))) |>
  tidy()|>
  mutate(p.value=as.character(p.value))|>
  kable( 
      caption = "Summary of Linear Model Coefficients", 
      col.names = c("Term", "Estimate", "Std. Error", "t value", "p value"))
```

```{r}
lm(formula = growth_rateY ~ growth_rateK + growth_rateL + growth_rateA + Pays, data = na.exclude(Dec)) |>
summary() |>
  (\(x) data.frame(
    Statistic = c("N", "R2", "Adjusted R2"),
    Value = c(x$df[1] + x$df[2], x$r.squared, x$adj.r.squared)
  ))() |>
  kable(
    caption = "Model Summary Statistics",
    col.names = c("Statistic", "Value")
  )

```

```{r}
# France
# Solow-Swan model
Solow <- Dec |>
  filter(Pays == "France") |>
  arrange(TIME_PERIOD) |>
  mutate(
    growth_rateY = (100 + growth_rateY) / 100,
    growth_rateK = (100 + growth_rateK) / 100,
    growth_rateL = (100 + growth_rateL) / 100,
    growth_rateA = (100 + growth_rateA) / 100
  ) |>
  filter(TIME_PERIOD >= 1990 & TIME_PERIOD <= 2015) |>
  mutate(
    growth_rate_totalL = 0.66*prod(growth_rateL),
    growth_rate_totalY = prod(growth_rateY),
    growth_rate_totalK = 0.33*prod(growth_rateK),
    growth_rate_totalA = 0.66*prod(growth_rateA)
  ) |>
  mutate(others = growth_rate_totalY - growth_rate_totalK - growth_rate_totalL - growth_rate_totalA)|>
  select(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others, growth_rate_totalY)
```

```{r}
#| output: false
# France 
Solow1 <- Solow |>
  pivot_longer(
    cols = c(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others),
    names_to = "Component",
    values_to = "Value"
  )

# Bar chart decomposition
graph5 <- ggplot(Solow1, aes(x = "", y = Value, fill = Component)) +
  geom_bar(stat = "identity") +
  labs(
    title = "Growth rate decomposition (France, Germany,USA, 1990-2015)",
    x = NULL,
    y = "Growth Rate Contribution",
    fill = "Component"
  ) +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
```


```{r}
# Germany
# Solow-Swan model
Solow2 <- Dec |>
  filter(Pays == "Allemagne") |>
  arrange(TIME_PERIOD) |>
  mutate(
    growth_rateY = (100 + growth_rateY) / 100,
    growth_rateK = (100 + growth_rateK) / 100,
    growth_rateL = (100 + growth_rateL) / 100,
    growth_rateA = (100 + growth_rateA) / 100
  ) |>
  filter(TIME_PERIOD >= 1990 & TIME_PERIOD <= 2015) |>
  mutate(
    growth_rate_totalL = 0.66*prod(growth_rateL),
    growth_rate_totalY = prod(growth_rateY),
    growth_rate_totalK = 0.33*prod(growth_rateK),
    growth_rate_totalA = 0.66*prod(growth_rateA)
  ) |>
  mutate(others = growth_rate_totalY - growth_rate_totalK - growth_rate_totalL - growth_rate_totalA)|>
  select(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others, growth_rate_totalY)
```

```{r}
#| output: false
# Germany 
Solow3 <- Solow2 |>
  pivot_longer(
    cols = c(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others),
    names_to = "Component",
    values_to = "Value"
  )

# Bar chart decomposition
graph6<-ggplot(Solow3, aes(x = "", y = Value, fill = Component)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(legend.position = "none", axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r}
# USA
# Solow-Swan model
Solow4 <- Dec |>
  filter(Pays == "États-Unis") |>
  arrange(TIME_PERIOD) |>
  mutate(
    growth_rateY = (100 + growth_rateY) / 100,
    growth_rateK = (100 + growth_rateK) / 100,
    growth_rateL = (100 + growth_rateL) / 100,
    growth_rateA = (100 + growth_rateA) / 100
  ) |>
  filter(TIME_PERIOD >= 1990 & TIME_PERIOD <= 2015) |>
  mutate(
    growth_rate_totalL = 0.66*prod(growth_rateL),
    growth_rate_totalY = prod(growth_rateY),
    growth_rate_totalK = 0.33*prod(growth_rateK),
    growth_rate_totalA = 0.66*prod(growth_rateA)
  ) |>
  mutate(others = growth_rate_totalY - growth_rate_totalK - growth_rate_totalL - growth_rate_totalA)|>
  select(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others, growth_rate_totalY)
```

```{r}
#| output: false
# USA 
Solow5 <- Solow4 |>
  pivot_longer(
    cols = c(growth_rate_totalL, growth_rate_totalA, growth_rate_totalK, others),
    names_to = "Component",
    values_to = "Value"
  )

# Bar chart decomposition
graph7 <-ggplot(Solow5, aes(x = "", y = Value, fill = Component)) +
  geom_bar(stat = "identity") +
  theme_minimal() +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

```{r}
#| fig.width: 12
#| fig.height: 4
graph5+graph6+graph7
```
$Y = K^\alpha (AL)^{1-\alpha}$
# Data sets descriptions

# Data analysis

# Conclusion
